---
title: "R Notebook"
output: html_notebook
---

Bring in packages
```{r}
library(readr)
library(tidyverse)
library (stats)
```

Set working directory and make parent csv and quarterly csv into dataframes
```{r}
setwd("C:/Users/maddi/Desktop/Pubs/Pub3-EconFlow/rstudio")

#ForR_AQ_Att <- read_csv("ForR_AQ_Att.csv")
QUART_AQ_and_Attendance_Closures <- read_csv("QUART_AQ_and_Attendance_Closures.csv")

#dfData_all<-data.frame(ForR_AQ_Att)
dfData_quart<-data.frame(QUART_AQ_and_Attendance_Closures)
```

Explore data with basic ggplots - one per gate
first: Arch Rock
```{r}
ggplot(data=dfData_quart, mapping=aes(x=QuarterYear, y=ArchRock, group=1))+geom_line()+ theme(axis.text.x=element_text(angle=-45, hjust=0.001))

```

second: Big Oak Flat
```{r}
ggplot(data=dfData_quart, mapping=aes(x=QuarterYear, y=BigOakFlat, group=1))+geom_line() + theme(axis.text.x=element_text(angle=-45, hjust=0.001))

```

third: Southgate
```{r}
ggplot(data=dfData_quart, mapping=aes(x=QuarterYear, y=Southgate, group=1))+geom_line()+ theme(axis.text.x=element_text(angle=-45, hjust=0.001))

```

fourth: Tioga Pass. Note very strong seasonality bc closures of pass every winter
```{r}
ggplot(data=dfData_quart, mapping=aes(x=QuarterYear, y=Southgate, group=1))+geom_line()+theme(axis.text.x=element_text(angle=-45, hjust=0.001))

```

Explore basic plot of unhealthy for sensitive 

```{r}
ggplot(data=dfData_quart, mapping=aes(x=QuarterYear, y=OverUnHS, group=1))+geom_line() + theme(axis.text.x=element_text(angle=-45, hjust=0.001))

```

Test correlation between smoky days and gate attendance (inelegant, but functional)
Data are only normal as individual years, not as full 20 year series

Full results: Each are moderately (>3) correlated, and the moderate correlation for all are statistically significant (p<0.05)
There is stronger correlation between visitor count and bad AQ days for TP because it isn't open in winter

WOrk on making this a for loop for elegance - in scratch section below
```{r}
corArch<-cor.test(QUART_AQ_and_Attendance_Closures$ArchRock, QUART_AQ_and_Attendance_Closures$OverUnHS, method=c("spearman"), exact = FALSE)
corBig<-cor.test(QUART_AQ_and_Attendance_Closures$BigOakFlat, QUART_AQ_and_Attendance_Closures$OverUnHS, method=c("spearman"), exact = FALSE)
corSouth<-cor.test(QUART_AQ_and_Attendance_Closures$Southgate, QUART_AQ_and_Attendance_Closures$OverUnHS, method=c("spearman"), exact = FALSE)
corTioga<-cor.test(QUART_AQ_and_Attendance_Closures$TiogaPass, QUART_AQ_and_Attendance_Closures$OverUnHS, method=c("spearman"), exact = FALSE)

print(corArch)
print(corBig)
print(corSouth)
print(corTioga)
```



Multiple Linear Regression for air quality and gates, per prospectus, then ANOVA w/o TP

I need to do this specifically for closure events
```{r}
fitAll<-lm(OverUnHS~ArchRock+BigOakFlat+Southgate+TiogaPass, data=dfData_quart)
fitNoTP<-lm(OverUnHS~ArchRock+BigOakFlat+Southgate, data=dfData_quart)

fitAll
fitNoTP


```

















SCRATCH
work on this to make it a for loop
```{r}
NUM_data<-as.numeric(unlist((QUART_AQ_and_Attendance_Closures)))


for (i in 6:length(NUM_data)) {
  a <- cor.test(NUM_data$OverUnHS, NUM_data[,i])
  print(paste(colnames(NUM_data)[i], " est:", a$estimate, " p=value:", a$p.value))
}
```

T-tests gate attendance to gate attendance
paired because they are NOT independent
Goals = high t, low p

Arch Rock test
I dont think this is doing what i think it is


```{r}
tArchToBig<-t.test(QUART_AQ_and_Attendance_Closures$ArchRock, QUART_AQ_and_Attendance_Closures$BigOakFlat, paired=TRUE)
tArchToSouth<-t.test(QUART_AQ_and_Attendance_Closures$ArchRock, QUART_AQ_and_Attendance_Closures$Southgate, paired=TRUE)
tArchToTioga<-t.test(QUART_AQ_and_Attendance_Closures$ArchRock, QUART_AQ_and_Attendance_Closures$TiogaPass, paired=TRUE)

print(tArchToBig)
print(tArchToSouth)
print(tArchToTioga)
```
